---
title: "Project1"
authors: "Caroline Fuccella, Mariam Guirguis, Caleb Neale, Nathan Weins"
output: html_notebook
---

# Hypotheses

## Derailments Cause more Damage

## Derailments Cause more Causalties

## Night time accidents cause more damage

### Load Data
```{r}
file.inputl <- function(my.path)
{
  my.dir <- getwd()
  setwd(my.path)
  my.files <- list.files(pattern=".csv")
  acts <- lapply(my.files,read.csv)
  setwd(my.dir)
  return(acts)
}

combine.data <- function(Data.List, Vars)
	{
		DF <- rbind(Data.List[[1]][, Vars])
		for(i in 2:length(Data.List))
	{
		DF <- rbind(DF, Data.List[[i]][, Vars])
	}
	DF
	}
```

```{r}
library(tidyverse)
files <- list.files("./TrainData")
oldwd <- getwd()
setwd("./TrainData")
list <- lapply(files, read.csv)
setwd(oldwd)
comvar <- intersect(colnames(list[[1]]), colnames(list[[8]]))
compiled <- combine.data(list, comvar)
```

### Bin by Night or not
```{r}
compiled <- compiled %>% mutate(is.night = if_else(
  (TIMEHR < 6 & AMPM == "AM") | (TIMEHR > 8 & AMPM == "PM"),
  1, 0)
  )
```

### Model Damage using only nighttime
```{r}
night.lm <- lm(ACCDMG~is.night, data = compiled)
summary(night.lm)
```
Positive coefficient value and significance of both the model and the predictor suggests that accidents at night do indeed have more damage than accidents at other times of day. 

### Plot damage by night or not
```{r}
library(plotly)
compiled %>% ggplot(aes(x=as.factor(is.night), y=ACCDMG)) + geom_boxplot()
compiled %>% ggplot(aes(x=as.factor(is.night), y=ACCDMG)) + geom_violin()
compiled %>% ggplot(aes(x=ACCDMG, color=as.factor(is.night))) + geom_density(alpha = 0.3) 
```
Difficult to see any visual differences due to large tails

```{r}
compiled %>% filter(ACCDMG < 1000000) %>% ggplot(aes(x=as.factor(is.night), y=ACCDMG)) + geom_boxplot()
compiled %>% filter(ACCDMG < 1000000) %>% ggplot(aes(x=as.factor(is.night), y=ACCDMG)) + geom_violin()
compiled %>% filter(ACCDMG < 1000000) %>% ggplot(aes(x=ACCDMG, color=as.factor(is.night))) + geom_density(alpha = 0.3) 
```
Still look quite similar, maybe higher median for nightime. 

### Model using time and interactions
```{r}
library(MASS)
full.time.lm <- lm(ACCDMG~(AMPM + TIMEHR + is.night)^2, data = compiled)
summary(full.time.lm)

anova(night.lm, full.time.lm)
```
We can then determine that at least one of the additional coefficients is nonzero.

```{r}
intermediate.lm <- lm(ACCDMG ~ is.night + TIMEHR:is.night + TIMEHR, data = compiled)
summary(intermediate.lm)
anova(night.lm, intermediate.lm)
anova(intermediate.lm, full.time.lm)
```
### Diagnostics for night only model
```{r}
plot(night.lm)
```
Massive nonnormal tail on the high end, likely transformation needed. No other anamaloies seen. 

### Transform data
```{r}
pos.data <- compiled %>% filter(ACCDMG > 0)
pos.night.lm <- lm(ACCDMG ~ is.night, pos.data)
bc <- boxcox(pos.night.lm)

lambda <- bc$x[which.max(bc$y)]

trans_model <- lm(((ACCDMG^lambda-1)/lambda) ~ is.night, pos.data)
plot(trans_model)
summary(trans_model)

```
Still not perfectly normal, but much better.

### Plot transformed data for better visual
```{r}
pos.data$ACCDMG.trans <- ((pos.data$ACCDMG^lambda-1)/lambda)
compiled %>% ggplot(aes(x=as.factor(is.night), y=ACCDMG.trans)) + geom_boxplot()
compiled %>% ggplot(aes(x=as.factor(is.night), y=ACCDMG.trans)) + geom_violin()
compiled %>% ggplot(aes(x=ACCDMG.trans, color=as.factor(is.night))) + geom_density(alpha = 0.3) 

```
Visual inspection shows no noticeable difference between night and not night. Small coefficient values in both the transformed and transformed model suggest that even though night is a statistically significant predictor, it may not be a practically useful one. 

### Interaction between speed and nighttime
```{r}
night.speed.lm <- lm(ACCDMG~as.factor(is.night) + TRNSPD + is.night:TRNSPD, compiled)
summary(night.speed.lm)
plot(night.speed.lm)

trans.night.speed.lm <- lm(((ACCDMG^lambda-1)/lambda)~as.factor(is.night) + TRNSPD + is.night:TRNSPD, pos.data)
summary(trans.night.speed.lm)
plot(trans.night.speed.lm)

```
### Interaction Plots

```{r}
interaction.plot(x.factor=compiled$TRNSPD, trace.factor = compiled$is.night, response=compiled$ACCDMG)

```


## Night time accidents cause more casualties